---
title: 'Statistical Inference Project: part 2'
output: html_document
---
**Instruction**: Now in the second portion of the class, we're going to analyze the ToothGrowth data in the R datasets package.  

1. Load the ToothGrowth data and perform some basic exploratory data analyses 
```{r loading, fig.height=3, fig.width=5, warning=F}
data(ToothGrowth); 
str(ToothGrowth); 

library(ggplot2)
ggplot(data=ToothGrowth, aes(x=factor(dose), y=len, color=supp))  + 
    geom_point()+ 
    xlab("Dose")+
    ylab("Length")+
    theme_bw()+
    ggtitle("Comparing tooth growth by supp and dose")
```

2. Provide a basic summary of the data.
*Summary:* 

- The data has 60 records and three variables(len, supp, dose). 
- There are three doses (0.5, 1, 2) under the dose varialble and 2 factors ("VC", "OJ") under the supp variable.
- The len variable has the tooth growth data, which can be divided into to six group according to the variables dose and supp. 
- Below is a table reshaped to show the mean and standard deviation of each group.
```{r, results='asis', echo=TRUE}

ToothGrowth$ID <-rep(1:10,6);library(reshape2); 
dat <- dcast(ToothGrowth, dose+supp~ID, value.var="len"); 

names(dat)[-(1:2)]<-paste("len_",names(dat[-(1:2)]), sep="");
dat$mean<-sapply(1:6, function(x) {mean(as.numeric(dat[x,3:12]))});
dat$sd<-sapply(1:6, function(x) {sd(as.numeric(dat[x,3:12]))});
dat$Group<-1:6
library(xtable);
print(xtable(dat[-(3:12)]),type='html')
```

3. Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose. (Use the techniques from class even if there's other approaches worth considering)

- calculate the 95% confidence interval between the two groups as stated in the table below. we use unpaired t-test assuming the groups are independent. 
- We first tried unequal variance t-test. we also tried t.test assuming equal variance. Both equal or unequal variances reach a same conclusion.

```{r CI, results='asis'}
conFun <- function(m=1, n=2, df=dat) {
    temp<-t.test(df[m,3:12],df[n,3:12],paired=FALSE, var.equal=FALSE)$conf[1:2];
    return(data.frame(
        Data_1 = paste("group_",m,": dose=",df[m,1],", supp=",df[m,2],sep=""),
        Data_2 = paste("group_",n,": dose=",df[n,1],", supp=",df[n,2],sep=""),
        Lowlimit = temp[1],
        Uplimit  = temp[2]))
    }

print(xtable(rbind(conFun(1,2),conFun(3,4),conFun(5,6))),type='html')
print(xtable(rbind(conFun(1,3),conFun(3,5),conFun(2,4),conFun(4,6))),type='html')
```


4. State your conclusions and the assumptions needed for your conclusions. 

- The 95% confidence intervals (or CI) between "OJ" and "VC" for dose =0.5 and 1 is above zero, however the 95% CI between "OJ" and "VC" for the dose of 2 includes zero. 
- We have 95% confidence that "OJ"" has more effect on the tooth growth that "VC" at lower dose (0.5 and 1).
- "OJ" and "VC" have similar effect on tooth growth at a higher dose of 2. 
